---
title: "create hist"
author: "Tourism Economics"
date: "Thursday, October 16, 2014"
output: 
  html_document: 
    keep_md: yes
---

Setup
```{r readchunk}
source('~/Project/R projects/hlt/scripts/functions.R')
```

Creates a us historical databank. Combines the STR data with selected macro data and calculates a few series
```{r load}

fpath <- c("~/Project/R projects/hlt/") 
#macro data
  load(paste(fpath,"output_data/out_macro_q.Rdata", sep=""))
# str data
  load(paste(fpath,"output_data/out_str_chn_q.Rdata", sep="")) 
  load(paste(fpath,"output_data/out_str_chn_m.Rdata", sep=""))
# GCT data
  load(paste(fpath,"output_data/out_gct_a.Rdata", sep="")) 
  load(paste(fpath,"output_data/out_gct_q.Rdata", sep="")) 
# Asia Cities data
  load(paste(fpath,"output_data/out_acities_a.Rdata", sep="")) 
  load(paste(fpath,"output_data/out_acities_q.Rdata", sep="")) 
# Global Cities data
  load(paste(fpath,"output_data/out_gcities_a.Rdata", sep="")) 
  load(paste(fpath,"output_data/out_gcities_q.Rdata", sep="")) 
# TDM data
  load(paste(fpath,"output_data/out_tdm_a.Rdata", sep="")) 
  load(paste(fpath,"output_data/out_tdm_q.Rdata", sep=""))
  
```

Decided to move the segment information from being bundled with the geography, to be part of the lodging variable. I guess the reason is that then I would have a clean geo column. 

```{r separate_seg}

str_q <- out_str_chn_q %>%
    data.frame(date=time(.), .) %>%
    gather(variable, value, -date) %>%
    # filter to drop the str_days variable
    filter(variable != "strdays") %>%
    # separate variable column
    separate(variable, c("seg", "variable"), 3) %>%
    separate(variable, c("geo", "variable"), sep="_") %>%
    # combine the seg and variable column
    mutate(variable = paste0(seg, variable)) %>%
    # create a single variable mnuemonic
    mutate(variable = paste(geo, variable, sep="_")) %>%
    select(-seg, -geo) %>%
    spread(variable, value) %>%
    read.zoo %>%
    as.xts
  
```


Calculate some real lodging series.
```{r calc real}
# to start, bring all series from the macro databank
# temp <- out_macro_q %>%
#   data.frame(date=time(.), .) %>%
#   read.zoo %>%
#   as.xts
# head(temp)

# merges dataframes. the all.=TRUE piece ensures all the rows
# in the first dataframe are included
#hist_q <- merge(temp, str_q, all=TRUE) 

  
######################
#
# converts some series to real
#

# first index the personal cons price deflator to average 100 in 2014
chn_pcindex <- index_q(out_macro_q$chn_pc, index_year=2014)
names(chn_pcindex) <- "chn_pcindex"
str_q <- merge(str_q, chn_pcindex)
autoplot(str_q$chn_pcindex)

# select the series that contain adr or revpar and convert to real
# the way this works is that matches works on a regular expression
# I wrote a regular expression that is taking columns with names that
# end with the text shown
# for reference on writing regular expressions, see
# http://www.regular-expressions.info/quickstart.html
real_df <- data.frame(str_q) %>%
  select(matches("adr$|adrsa$|revpar$|revparsa$")) %>% 
  mutate_each(funs(ind = ( . / chn_pcindex)*100))
# adds on a time index to get it back to xts
temp <- data.frame(date=time(str_q)) 
real_df <- cbind(temp,real_df)
real <- read.zoo(real_df)
real <- xts(real)

# renames series 
tempnames <- names(real)
tempnames <- paste(tempnames,"r",sep="")
tempnames
names(real) <- tempnames
rm(tempnames)

autoplot(window(real$chn_upaadrsar, start="2000-01-01", end="2015-10-01"))
autoplot(window(str_q$chn_upaadrsa, start="2000-01-01", end="2015-10-01"))

# merges onto ushist_q
str_q <- merge(str_q, real)
autoplot(window(str_q$chn_upaadrsar, start="2000-01-01", end="2015-10-01"))
```

Looking at what's in quarterly databank
```{r look_q}
# which segments or markets are in the data frame, just for observation
# not used anywhere
a <- grep(pattern="demt$", colnames(str_q), value=TRUE)
a
a <- gsub(pattern="demt$",replacement="",a)
a

b <- grep(pattern="chn_", colnames(str_q), value=TRUE)
b
```

Creating quarterly dataframe
```{r create_q}

str_qt <- str_q %>%
  data.frame(date=time(.), .) %>%
  gather(variable, value, -date) %>%
  separate(variable, c("area_sh", "variable"), sep = "_") %>%
  spread(variable, value)

# The challenge here is that I have a lodging dataframe, with a mix of cities
# and a country. Then I have a macro dataframe with just China. And then 
# I have some city dataframes that have both, and then I have the tdm
# dataframe with just China. Do I try to put them all together? Why?
# As part of the challenge, GDP for example, is in several of the dataframes
# so when I joined, I started ending up with several columns of GDP.
# My initial approach was to be careful. I was going to set up dataframes
# of what I'm taking from each source. With the goal of taking each concept
# only once for each geography, and then I'd figure out how to combine them

# Then I realized an approach of adding a source column to each source, 
# thencombining everything, and then filtering 
# to drop certain items, or keep only certain items, and then checking 
# what I had in the end. 

# take everything from the tidy str dataframe
take_str_q <- str_qt %>%
    gather(variable, value, -date, -area_sh) %>%
  mutate(src = "str")

# take everything from the macro dataframe
take_macro_q <- out_macro_q %>%
  data.frame(date=time(.), .) %>%
  gather(variable, value, -date) %>%
  separate(variable, c("area_sh", "variable"), sep = "_") %>%
  mutate(src = "macro")

# take GDP for some cities from Asia cities
take_acities_q <- out_acities_q %>%
  select(date, area_sh, gdp) %>%
  gather(variable, value, -date, -area_sh) %>%
  mutate(src = "acities")

# take GDP for other cities form Global cities
take_gcities_q <- out_gcities_q %>%
  select(date, area_sh, gdp) %>%
  gather(variable, value, -date, -area_sh) %>%
  mutate(src = "gcities")

# take everything from tdm
take_tdm_q <- out_tdm_q %>%
  gather(variable, value, -date, -area_sh) %>%
  mutate(src = "tdm")

# combine everything
temp_a <- bind_rows(take_str_q, take_macro_q, take_acities_q, take_gcities_q, take_tdm_q) %>%
  filter(!is.na(value)) 

# filter to remove certain rows that I know are being duplicated based on my sources
temp_b <- temp_a %>%
  # sort of amazing, but this gives me all records that don't meet
  # the condition of having gcities as a source and not being in the 
  # list. So of those with gcities as a source, I'm only keeping
  # chgduchn
  filter(!(src == "gcities" & !area_sh %in% c("chgduchn")))
  # the following was evidently not necessary
  #filter(!(src == "acities" & !area_sh %in% c("beijichn", "gngzhchn", "shangchn"))) %>%

# as long as check_dup is empty there are evidently no duplicates
# based partly on the following link, I removed the negative 
#https://www.quora.com/How-do-I-remove-entire-rows-of-duplicates-based-on-two-different-columns-in-a-R-dataframe
check_dup <- temp_b[duplicated(temp_b[,c('date', 'area_sh', 'variable')]),]


hist_q <- temp_b %>%
  select(-src) %>%
  spread(date, value) %>%
  mutate(indicator_code = paste(area_sh, variable, sep = "_")) %>%
  select(area_sh, variable, indicator_code, everything())

```




Create annual STR data
```{r create_str_a}

# put quarterly dataframe back into a zoo form
temp_q <- hist_q %>%
  select(-area_sh, -variable) %>%
  gather(date, value, -indicator_code) %>%
  spread(indicator_code, value) %>%
  data.frame() %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  as.xts

###########
#
# start by creating an annual version of lodging data

# select series that should be converted to annual by summing
# I wrote a regular expression that is looking for certain text strings
# for reference on writing regular expressions, see
# http://www.regular-expressions.info/quickstart.html
suma <- data.frame(temp_q) %>%
  select(matches("demt$|supt$|rmrevt$")) %>%
  as.xts()

# not sure why I needed this. Maybe it didn't like that it ended on a first quarter
suma <- window(suma, start = as.POSIXct("1980-01-01"), end = as.POSIXct("2029-10-01"))

# this function is one I defined, it converts all the columns in 
# an xts object to annual. Must be an xts object to start with
suma <- q_to_a_xts(suma, type="sum")

# takes the summed data and spreads it into a tidy format with
# tidyr and then calculates the occupancy and revpar series
# first needs to go from xts to dataframe
tb2 <- data.frame(date=time(suma), suma)%>% 
  # creates column called segvar that contains the column names, and one next to 
  # it with the values, dropping the time column
  gather(segvar, value, -date, na.rm = FALSE) %>%
  # in the following the ^ means anything not in the list
  # with the list being all characters and numbers
  # so it separates segvar into two colums using sep
  separate(segvar, c("area_sh", "variable"), sep = "[^[:alnum:]]+") %>%
  # added for China, separates off the chain scale segment
  separate(variable, c("seg", "variable"), sep = 3) %>%
  # keeps seg as a column and spreads variable into multiple columns containing
  # containint the values
  spread(variable,value) %>%
  # adds new calculated column
  mutate(occ = demt / supt) %>%
  # adds another column
  mutate(revpar = rmrevt / supt) %>%
  mutate(adr = rmrevt / demt)

# takes it from a tidy format, recreates the variable mnuemonic,
# and then creates the unique
# variable names and then reads into a zoo object spliting on the 
# second column
str_at  <- tb2 %>%
  gather(variable, value, -date, -area_sh,-seg) %>%
  mutate(variable = paste0(seg, variable)) %>%
  select(-seg)

```

Create annual databank
```{r create_a}

# take everything from the tidy str dataframe
take_str_a <- str_at %>%
  mutate(src = "str") %>%
  mutate(variable = as.character(variable))

# take everything from the macro dataframe - doesn't exist would require work
# take_macro_a <- out_macro_a %>%
#   data.frame(date=time(.), .) %>%
#   gather(variable, value, -date) %>%
#   separate(variable, c("area_sh", "variable"), sep = "_") %>%
#   mutate(src = "macro")

# take GDP for some cities from Asia cities
take_acities_a <- out_acities_a %>%
  select(date, area_sh, gdp) %>%
  gather(variable, value, -date, -area_sh) %>%
  mutate(src = "acities") %>%
  mutate(variable = as.character(variable))

# take GDP for other cities form Global cities
take_gcities_a <- out_gcities_a %>%
  select(date, area_sh, gdp) %>%
  gather(variable, value, -date, -area_sh) %>%
  mutate(src = "gcities") %>%
  mutate(variable = as.character(variable))

# take everything from tdm
take_tdm_a <- out_tdm_a %>%
  gather(variable, value, -date, -area_sh) %>%
  mutate(src = "tdm") %>%
  mutate(variable = as.character(variable))

# combine everything
temp_a <- bind_rows(take_str_a, take_acities_a, take_gcities_a, take_tdm_a) %>%
  filter(!is.na(value)) 


# filter to remove certain rows that I know are being duplicated based on my sources
temp_b <- temp_a %>%
  # sort of amazing, but this gives me all records that don't meet
  # the condition of having gcities as a source and not being in the 
  # list. So of those with gcities as a source, I'm only keeping
  # chgduchn
  filter(!(src == "gcities" & !area_sh %in% c("chgduchn")))
  # the following was evidently not necessary
  #filter(!(src == "acities" & !area_sh %in% c("beijichn", "gngzhchn", "shangchn"))) %>%

# as long as check_dup is empty there are evidently no duplicates
# based partly on the following link, I removed the negative 
#https://www.quora.com/How-do-I-remove-entire-rows-of-duplicates-based-on-two-different-columns-in-a-R-dataframe
check_dup <- temp_b[duplicated(temp_b[,c('date', 'area_sh', 'variable')]),]

hist_a <- temp_b %>%
  select(-src) %>%
  spread(date, value) %>%
  mutate(indicator_code = paste(area_sh, variable, sep = "_")) %>%
  select(area_sh, variable, indicator_code, everything())

```

Creating monthly historical databank
```{r create_m}
# not that much that needs to be done
hist_m <- out_str_chn_m

```

### Writing outputs
```{r write_output, echo=FALSE}
# quarterly
  # writes csv versions of the output files
  write.zoo(hist_q, file=paste(fpath, "output_data/hist_q.csv", sep=""), sep=",")
  # saves Rdata versions of the output files
  save(hist_q, file=paste(fpath, "output_data/hist_q.Rdata", sep=""))

# monthly
  # writes csv versions of the output files
  write.zoo(hist_m, file=paste(fpath,"output_data/hist_m.csv",sep=""), sep=",")
  # saves Rdata versions of the output files
  save(hist_m, file=paste(fpath,"output_data/hist_m.Rdata", sep=""))

# annual
  # writes csv versions of the output files
  write.zoo(hist_a, file=paste(fpath,"output_data/hist_a.csv",sep=""), sep=",")
  # saves Rdata versions of the output files
  save(hist_a, file=paste(fpath,"output_data/hist_a.Rdata", sep=""))
```
